Precise temporal localisation of M/EEG effects with Bayesian generalised additive multilevel models

### 2025-04-10 brms_meeg.tex
pushing a cleaner version of the intro

\documentclass[
  doc,
  floatsintext,
  longtable,
  a4paper,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

@@

\geometry{inner=1in, outer=1in}
\fancyhfoffset[LE,RO]{0cm}

@@

\abstract{Time-resolved electrophysiological measurements such as those
obtained through magneto- or electro-encephalography (M/EEG) offer a
unique window into the neural activity underlying cognitive processes.
Researchers are often interested in determining whether and when these
signals differ across experimental conditions or participant groups. The
conventional approach involves mass-univariate statistical testing
across time and space, followed by corrections for multiple comparisons
such as cluster-based inference. While effective for controlling error
rates at the cluster-level, cluster-based inference comes with a
significant limitation: by shifting the focus of inference from
individual time points to clusters, it makes difficult to draw precise
conclusions about the onset or offset of observed effects. Here, we
introduce a \emph{model-based} approach for analysing M/EEG timeseries
such as event-related potentials (ERPs) or decoding performance over
time. Our method leverages Bayesian generalised additive multilevel
models (GAMMs), providing posterior probabilities that an effect is
above zero (or above chance) at each time point, while naturally
accounting for temporal dependencies and between-subject variability.
Using both simulated and actual M/EEG datasets, we demonstrate that this
approach substantially outperforms conventional methods in estimating
the onset and offset of neural effects, yielding more precise and
reliable results. We provide an R package implementing the method and
show how it can be seamlessly integrated into M/EEG analysis pipelines
using MNE-Python.}

@@

The conventional approach involves mass-univariate statistical testing
with a unique threshold, followed by corrections for multiple
comparisons with the goal of maintaining the familywise error rate
(FWER) or false discovery rate (FDR) at the nominal level. Cluster-based
inference is the most common way of achieving this sort of error control
in the M/EEG literature, being the default option in several software
programs. While effective for controlling error rates, cluster-based
inference comes with a significant limitation: they shift the focus of
inference from individual time points to clusters, making it difficult
to draw precise conclusions about the onset or offset of observed
effects (\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). As
pointed by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007}); ``there
is a conflict between this interest in localized effects and our choice
for a global null hypothesis: by controlling the FA {[}false alarm{]}
rate under this global null hypothesis, one cannot quantify the
uncertainty in the spatiotemporal localization of the effect''. Even
worse, as noted by Rosenblatt et al.
(\citeproc{ref-rosenblatt2018}{2018}), cluster-based inference suffers
from low spatial resolution: ``Since discovering a cluster means that
`there exists at least one voxel with an evoked response in the
cluster', and not that `all the voxels in the cluster have an evoked
response', it follows that the larger the detected cluster, the less
information we have on the location of the activation.'' As a
consequence, cluster-based inferenc has been shown\ldots{}

To overcome the limitations of cluster-based inference, we introduce a
novel model-based approach for precisely localising M/EEG effects in
time (space, and other dimensions). The proposed approach, based on
Bayesian generalised additive multilevel models, allows quantifying the
posterior probability of effects being above chance at the level of
timesteps, sensors, voxels, etc., while naturally taking into account
spatiotemporal dependencies present in M/EEG timeseries. We compare the
performance of the proposed approach to well-established alternative
methods using both simulated and actual M/EEG data and show that it
significantly outperforms alternative methods in estimating the onset
and offset of M/EEG effects.

@@

Different methods exist to control the family-wiser error rate (FWER),
defined as the type-1 error rate (false positive) over an ensemble
(family) of tests\ldots{} for instance the Bonferroni correction
(\citeproc{ref-dunn1961}{Dunn, 1961}), however this method is generally
overconservative as it assumes statistical independence of tests, an
assumptions that is clearly violated in the context of M/EEG timeseries
with spatiotemporal dependencies\ldots{} or the false discovery rate
(FDR), defined as the proportion of false positive among positive tests
(e.g., \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
\citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001})\ldots{}

@@

where \(h_{0}\) is typically \(0\) and parameters \(E\) and \(H\) are
set a priori (typically to 0.5 and 2, respectively) and control the
influence of the extend and height on the TFCE. Then, p-value for
timestep/voxel \(t\) is computed by comparing it TFCE with the null
distribution of TFCE values. For each permuted signal, we keep the
maximal value over the whole signal for the null distribution of the
TFCE. The TFCE combined with permutation (assuming a large enough number
of permutations) has been shown to provide accurate type 1 FWER (e.g.,
\citeproc{ref-pernet2015}{Pernet et al., 2015}). However, previous
simulation work showed that cluster-based methods have indeed poor
performance in localising the onset of M/EEG effects (e.g.,
\citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).

To sum up this section, cluster-based inference main limitations is that
it provides inference at the cluster level only, not allowing inference
at the level of timesteps, sensors, etc. Thus, it does not allow
inferring the localisation of effects (\citeproc{ref-maris2011}{Maris,
2011}; for more details on cluster-based inference, see for instance
\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). In the
following, we briefly review previous modelling work. Then, we provide a
brief introduction to Bayesian statistical modelling, before moving to
generalised additive models (GAMs) and generalised additive multilevel
models (GAMMs) to illustrate how they can be used to precisely localise
the onset and offset of M/EEG effects.

@@

Disentangling overlapping processes
(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
Weighting single trials (\citeproc{ref-pernet2022}{Pernet,
2022})\ldots{} The LIMO toolbox (\citeproc{ref-pernet2011}{Pernet et
al., 2011})\ldots{}

@@

\CommentTok{\# fitting the BGAM}
\NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# cubic regression splines with k{-}1 basis functions}
    \CommentTok{\# eeg \textasciitilde{} condition + s(time, bs = "cr", k = 20, by = condition),}
    \CommentTok{\# thin{-}plate regression splines with k{-}1 basis functions}
\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
    \AttributeTok{data =}\NormalTok{ ppt\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

However, the previous model only included constant (fixed) effects, thus
not properly accounting for between-participant variability. We next fit
a multilevel version of the BGAM (BGAMM). Although it is possible to fit
a BGAMM at the single-trial level, we present a resource-efficient
version of the model that is fitted directly on the by-participant
summary statistics (mean and SD), similar to what is done in
meta-analysis.

@@

We depict the posterior predictions together with the posterior estimate
of the slope for \texttt{condition} at each timestep
(Figure~\ref{fig-plot-post-slope}). This figure suggests that the BGAMM
provides an adequate description of the simulated data.

@@

We then compute the posterior probability of the slope for
\texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
left). We can also express this as the ratio of posterior probabilities
(i.e., \(p/(1-p)\)) and visualise the timecourse of this ratio
superimposed with the conventional thresholds on evidence ratios
(Figure~\ref{fig-post-prob-ratio}, right). Note that a ratio of 10 means
that the probability of the difference being above 0 is 10 times higher
than the probability of the difference not being above 0, given the
data, the priors, and other model's assumptions. Thresholding the
posterior probability ratio thus provides a model-based approach for
estimating the onset and offset of M/EEG effects.

@@

\caption{\label{fig-post-prob-ratio}Left: Posterior probability of the
EEG difference (slope) being above 0 according to the BGAMM. Right:
Ratio of posterior probability according to the BGAMM (on a log10
scale). Timesteps above threshold (10) are highlighted in green.}

@@

\subsection{Comparing the identified onsets/offsets to other
approaches}\label{comparing-the-identified-onsetsoffsets-to-other-approaches}

@@

To assess the accuracy of group-level onset estimation, the various
methods were compared using the bias (i.e., median(estimated-true)),
median absolute error (MAE), root mean square error (RMSE), variance,
and median absolute deviation (MAD) of onset/offset estimates computed
on 1000 simulated datasets. As in Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), each participant was
assigned a random onset betwee, 150 and 170ms.

@@

We assessed the reliability of the proposed approach using a form of
permutation-based split-half reliability (as for instance in
\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}), which consisted
of the following steps. First, we created 1000 split halves of the data
(i.e., with half the participants in the original data, that is, 16
participants). For each split, we estimated the onset/offset using all
methods described previously. Third, we summarised the distribution of
onset/offset estimates using the median ``error'' (i.e., difference
between the split estimate and the estimate obtained using the full
dataset) and the variance across splits. This approach allows assessing
how similar the estimate of each half split is to the full dataset (thus
acting as a proxy for the population) and how variable the estimates are
(across split halves).

@@

{\caption{{Summary statistics of the onset and offset estimates for each
method (ordered by MAE).}{\label{tbl-simulation-results}}}}

@@

Figure~\ref{fig-reliability} shows the median difference between the
onset and offset estimates from each data split and the onset and offset
estimates from the full dataset (x-axis) along with the variance of its
onset and offset estimates across data splits (error bar). This figure
reveals that the \texttt{BGAM} \emph{onset and offset} estimates on each
split are the closest to the estimates from the full dataset on average
(0ms difference for the onset estimate and 5ms difference for the offset
estimate). The \texttt{Raw\ p-value} method has similar performance, but
given the aberrant estimates it produces (cf.
Figure~\ref{fig-onset-offset}), the fact that it is consistent between
data splits and the full dataset is not convincing on its own. The
\texttt{Change\ point} method also has a very good performance (i.e.,
very low difference between split estimates and full estimates), but
produces too short cluster of significant decoding performance (cf.
Figure~\ref{fig-onset-offset}).\footnote{As in Rousselet
  (\citeproc{ref-rousselet_using_2025}{2025}), we fixed the number of
  expected change points to two in the binary segmentation algorithm,
  thus producing always one cluster.} Overall, the figure reveals that
for all other methods, split datasets produce later onset estimates and
earlier offset estimates (as compared to the estimates from the model
fitted on the full dataset). These results highlight that the results of
any method should be assessed using both i) good asymptotic properties
on simulated data, ii) sensible identified clusters in actual data, and
iii) reliable/stable estimates on actual data.

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GAM with two temporal dimensions}
\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# 2D thin{-}plate spline (tp)}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 30),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
    \CommentTok{\# file = "models/timegen\_gam\_t2\_k30.rds"}
\NormalTok{    )}

>>>>

\documentclass[
  man,
  floatsintext,
  longtable,
  a4paper,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

@@

@@

\abstract{Time-resolved electrophysiological measurements such as those
obtained through magneto- or electro-encephalography (M/EEG) offer a
unique window into the neural activity underlying cognitive processes.
Researchers are often interested in determining whether and when these
signals differ across experimental conditions or participant groups. The
conventional approach involves mass-univariate statistical testing
across time and space, followed by corrections for multiple comparisons
such as cluster-based inference. While effective for controlling error
rates at the cluster-level, cluster-based inference comes with a
significant limitation: by shifting the focus of inference from
individual time points to clusters, it makes difficult to draw precise
conclusions about the onset or offset of observed effects. Here, we
introduce a \emph{model-based} approach for analysing M/EEG timeseries
such as event-related potentials (ERPs) or decoding performance over
time. Our approach leverages Bayesian generalised additive multilevel
models, providing posterior probabilities that an effect is above zero
(or above chance) at each time point, while naturally accounting for
temporal dependencies and between-subject variability. Using both
simulated and actual M/EEG datasets, we demonstrate that this approach
substantially outperforms conventional methods in estimating the onset
and offset of neural effects, yielding more precise and reliable
results. We provide an R package implementing the method and describe
how it can be integrated into M/EEG analysis pipelines using
MNE-Python.}

@@

The conventional approach involves mass-univariate statistical testing
through time and/or space followed by some form or correction for
multiple comparisons with the goal of maintaining the familywise error
rate (FWER) or false discovery rate (FDR) at the nominal level (e.g.,
5\%). Cluster-based inference is the most common way of achieving this
sort of error control in the M/EEG literature, being the recommended
approach in several software programs (e.g., \texttt{EEGlab} or
\texttt{MNE-Python}, \citeproc{ref-delorme2004}{Delorme \& Makeig,
2004}; \citeproc{ref-gramfort2013}{Gramfort, 2013}). While effective for
controlling error rates, cluster-based inference comes with a
significant limitation: by shifting the focus of inference from
individual time points to clusters, it prevents the ability to draw
precise conclusions about the onset or offset of observed effects
(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). As
pointed out by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007});
``there is a conflict between this interest in localized effects and our
choice for a global null hypothesis: by controlling the FA {[}false
alarm{]} rate under this global null hypothesis, one cannot quantify the
uncertainty in the spatiotemporal localization of the effect''. Even
worse, Rosenblatt et al. (\citeproc{ref-rosenblatt2018}{2018}) note that
cluster-based inference suffers from low spatial resolution: ``Since
discovering a cluster means that `there exists at least one voxel with
an evoked response in the cluster', and not that `all the voxels in the
cluster have an evoked response', it follows that the larger the
detected cluster, the less information we have on the location of the
activation.'' As a consequence, cluster-based inference is expected to
perform poorly for localising the onset of M/EEG effects; a property
that was later demonstrated in simulations studies (e.g.,
\citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).

To overcome the limitations of cluster-based inference, we introduce a
novel \emph{model-based} approach for precisely localising M/EEG effects
in time, space, and other dimensions. The proposed approach, based on
Bayesian generalised additive multilevel models, allows quantifying the
posterior probability of effects being above chance at the level of
timesteps, sensors, voxels, or other dimensions, while naturally taking
into account spatiotemporal dependencies present in M/EEG timeseries. We
compare the performance of the proposed approach to well-established
alternative methods using both simulated and actual M/EEG data and show
that it significantly outperforms alternative methods in estimating the
onset and offset of M/EEG effects.

@@

The issue with multiple comparisons\ldots{} Different methods exist to
control the family-wiser error rate (FWER), defined as the type-1 error
rate (false positive) over an ensemble (family) of tests\ldots{} for
instance the Bonferroni correction (\citeproc{ref-dunn1961}{Dunn,
1961}), however this method is generally overconservative as it assumes
statistical independence of tests, an assumptions that is clearly
violated in the context of M/EEG timeseries with spatiotemporal
dependencies\ldots{} or the false discovery rate (FDR), defined as the
proportion of false positive among positive tests (e.g.,
\citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
\citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001})\ldots{}

@@

where \(h_{0}\) is typically \(0\) and parameters \(E\) and \(H\) are
set a priori (typically to 0.5 and 2, respectively) and control the
influence of the extend and height on the TFCE. Note that in practise,
this intergral is approximated by a sum over small \(h\) increments.
Then, p-value for timestep/voxel \(t\) is computed by comparing it TFCE
with the null distribution of TFCE values. For each permuted signal, we
keep the maximal value over the whole signal for the null distribution
of the TFCE. The TFCE combined with permutation (assuming a large enough
number of permutations) has been shown to provide accurate type 1 FWER
(e.g., \citeproc{ref-pernet2015}{Pernet et al., 2015}). However,
previous simulation work showed that cluster-based methods (including
the TFCE method) perform poorly in localising the onset of M/EEG effects
(e.g., \citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).

To sum up this section, cluster-based inference main limitations is that
it provides inference at the cluster level only, not allowing inference
at the level of timesteps, sensors, etc. As a consequence, it does not
allow inferring the localisation of effects (for more details on
cluster-based inference, see for instance
\citeproc{ref-maris2011}{Maris, 2011}; \citeproc{ref-maris2007}{Maris \&
Oostenveld, 2007}; \citeproc{ref-sassenhagen2019}{Sassenhagen \&
Draschkow, 2019}). In the following, we briefly review previous M/EEG
modelling work. Then, we provide a brief introduction to generalised
additive models (GAMs) and Bayesian generalised additive multilevel
models (BGAMMs) to illustrate how these models can be used to precisely
localise the onset and offset of M/EEG effects.

@@

Disentangling overlapping processes
(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
Weighting single trials (\citeproc{ref-pernet2022}{Pernet,
2022})\ldots{} The LIMO toolbox (\citeproc{ref-pernet2011}{Pernet et
al., 2011}) using linear and 2-stage regression (not a proper multilevel
model)\ldots{}

@@

\CommentTok{\# fitting the BGAM}
\NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# thin{-}plate regression splines with k{-}1 basis functions}
\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
    \AttributeTok{data =}\NormalTok{ ppt\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

However, the previous model only included constant (fixed) effects, thus
not properly accounting for between-participant variability. We next fit
a multilevel version of the BGAM (BGAMM, see
\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). Although it is
possible to fit a BGAMM at the single-trial level, we present a
computationally-lighter version of the model that is fitted directly on
by-participant summary statistics (mean and SD), similar to what is done
in meta-analysis.

@@

We depict the posterior predictions together with the posterior estimate
of the slope for \texttt{condition} at each timestep
(Figure~\ref{fig-plot-post-slope}). This figure suggests that the BGAMM
provides an adequate description of the simulated data (see further
posterior predictive checks in Section~\ref{sec-basis}).

@@

We then compute the posterior probability of the slope for
\texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
left). From this quantity, we then compute the ratio of posterior
probabilities (i.e., \(p/(1-p)\)) and visualise the timecourse of this
ratio superimposed with the conventional thresholds on evidence ratios
(Figure~\ref{fig-post-prob-ratio}, right). Note that a ratio of 10 means
that the probability of the difference being above 0 is 10 times higher
than the probability of the difference not being above 0, given the
data, the priors, and other model's assumptions. Thresholding the
posterior probability ratio thus provides a model-based approach for
estimating the onset and offset of M/EEG effects.

@@

\caption{\label{fig-post-prob-ratio}Left: Posterior probability of the
EEG difference (slope) being above 0 according to the BGAMM. Right:
Ratio of posterior probability according to the BGAMM (on a log10
scale). Timesteps above threshold (10) are highlighted in green. NB: the
minimum and maximum possible ratio values are determined (bounded) by
the number of available posterior samples in the model.}

@@

\subsection{Comparing the onsets/offsets estimates from other
approaches}\label{comparing-the-onsetsoffsets-estimates-from-other-approaches}

@@

To assess the accuracy of group-level onset estimation, the various
methods were compared using the bias (i.e., median(estimated-true)),
median absolute error (MAE), root mean square error (RMSE), variance,
and median absolute deviation (MAD) of onset/offset estimates computed
on 1,000 simulated datasets. As in Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), each participant was
assigned a random onset between 150 and 170ms.

@@

We assessed the reliability of the proposed approach using a form of
permutation-based split-half reliability (as for instance in
\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}), which consisted
of the following steps. First, we created 1,000 split halves of the data
(i.e., with half the participants in the original data, that is, 16
participants). For each split, we estimated the onset/offset using all
methods described previously. Third, we summarised the distribution of
onset/offset estimates using the median ``error'' (i.e., difference
between the split estimate and the estimate obtained using the full
dataset) and the variance across splits. This approach allows assessing
how similar the estimate of each half split is to the full dataset (thus
acting as a proxy for the population) and how variable the estimates are
across split halves.

@@

{\caption{{Summary statistics of the onset and offset estimates for each
method (ordered by MAE).}{\label{tbl-simulation-results}}}
\vspace{-20pt}}

@@

Figure~\ref{fig-reliability} shows the median difference between the
onset and offset estimates from each data split and the onset and offset
estimates from the full dataset (x-axis) along with the variance of its
onset and offset estimates across data splits (error bar). This figure
reveals that the \texttt{BGAM} \emph{onset and offset} estimates on each
split are the closest to the estimates from the full dataset on average
(0ms difference for the onset estimate and 5ms difference for the offset
estimate). The \texttt{Raw\ p-value} method has similar performance, but
given the aberrant estimates it produces (cf.
Figure~\ref{fig-onset-offset}), the fact that it is consistent between
data splits and the full dataset is not convincing on its own. The
\texttt{Change\ point} method also has a very good performance (i.e.,
very low difference between split estimates and full estimates), but
produces too short cluster of significant decoding performance (cf.
Figure~\ref{fig-onset-offset}).\footnote{As in Rousselet
  (\citeproc{ref-rousselet_using_2025}{2025}), we fixed the number of
  expected change points to two in the binary segmentation algorithm,
  thus producing always one cluster.} Overall, the figure reveals that
for all other methods, split datasets produce later onset estimates and
earlier offset estimates (as compared to the estimates from the model
fitted on the full dataset). These results highlight some desirable
properties for a method aiming to precisely and reliably estimate the
onset and offset of M/EEG effects, namely, it should i) have good
asymptotic properties on simulated data, ii) provide sensible identified
clusters in actual data, and iii) provide reliable/stable estimates on
actual data.

@@

\bibitem[\citeproctext]{ref-delorme2004}
Delorme, A., \& Makeig, S. (2004). EEGLAB: an open source toolbox for
analysis of single-trial EEG dynamics including independent component
analysis. \emph{Journal of Neuroscience Methods}, \emph{134}(1), 9--21.
\url{https://doi.org/10.1016/j.jneumeth.2003.10.009}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GAM with two temporal dimensions}
\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# 2D thin{-}plate spline (tp)}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"} \CommentTok{\# k = 10}
    \CommentTok{\# file = "models/timegen\_gam\_t2\_k20.rds" \# k = 20}
\NormalTok{    )}
